---
title: "TidyVerse Tutorial"
author: "Smaranya Dey"
date: "6/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Tidyverse allows the R data frame to be a specific data format _tidy_. For more information on tidyverse, refer to their official website: <https://www.tidyverse.org/packages/>. 

Let us start with loading the package.

```{r}
library(tidyverse)
```

As we can see *tidyverse* includes a bunch of different packages like *ggplot2*, *purrr*, *dplyr*, *tidyr* and many more. 

To know and get familiar with *tidyverse*, we will use a few datasets like _murders_ from _dslabs_ package, _mtcars_, _nycflights_ etc.

### 1. dplyr package

We will get our hands dirty with the _dplyr_ package first. The definition of this package, as given in <https://dplyr.tidyverse.org>, is that, _dplyr_ is the grammar of data manipulation, providing a consistent set of verbs that help you solve the most common data manipulation challenges. Some of the functions widely used in _dplyr_ package are as follows:

* _mutate()_: Adding new variables that are functions of the existing variables

* _select()_: picks variables with their names

* _filter()_: picks cases based oon their values

* _summarise()_: reduces multiple values to single summary

* _arrange()_: changes ordering of the rows

* _group_by(): allows you to perform any operation by 'group'

We are going to follow *Chapter 3: Data Tranformation with dplyr* from *R for Data Science* book, the website <https://rafalab.github.io/dsbook/tidyverse.html> to exercise different functions from the package. They follow _nycflights13_ dataset and _murders_ dataset from _dslabs_ package. So, we are going to install and load the packages here.

```{r}
library(dslabs)
library(nycflights13)
```



